---
title: "BIS015L Project"
authors: Chloe, Emily, Ibrahim
date: "`r Sys.Date()`"
output:
  html_document: 

    keep_md: yes

---

## Load Libraries 


```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(janitor)
library(svglite)
library(shiny)
library(shinydashboard)
library(paletteer)
options(scipen=999)
``` 

## Load the Data
```{r}
disease <- readr::read_csv("Data/infectious-diseases-by-county-year-and-sex 2.csv")
disease
```

## Tidy Names of Data
```{r}
disease_tidy <- janitor::clean_names(disease)
names(disease_tidy)
```

## Remove data we will not use
```{r}
disease_data <- disease_tidy %>% 
  select(disease, county, year, sex, cases, population)
disease_data
```

## Summary of Data

```{r}
skimr::skim(disease_data)
```

```{r}
dim(disease_data)
```

```{r}
naniar::miss_var_summary(disease_data)
```

## General Data Analysis

Must first convert cases and population to numeric.
```{r}
disease_data$cases <- as.numeric(disease_data$cases) 
disease_data$population <- as.numeric(disease_data$population) 
```

1. Which disease is most prevalent throughout all of California?
```{r}
disease_data %>%
  filter(county!="California", sex=="Total") %>%
  group_by(disease) %>%
  summarise(total_cases = sum(cases)) %>%
  arrange(desc(total_cases))
```
```{r}
disease_data %>%
  filter(county!="California", disease == "Campylobacteriosis", sex == "Total") %>%
  summarise(total_cases = sum(cases))
```
*Campylobacteriosis is the most prevalent disease in California. It is an infection caused by consuming any raw poultry or meat.*

2. What are the top 5 most prevalent diseases in California?

```{r}
top_5_diseases <- disease_data %>%
  filter(county!="California", sex=="Total") %>%
  group_by(disease) %>%
  summarise(total_cases = sum(cases)) %>%
  top_n(5) %>%
  arrange(desc(total_cases))
top_5_diseases
```

```{r}
palette <- paletteer_d("wesanderson::Moonrise3")

top_5_diseases %>%
  ggplot(aes(x=reorder(disease, total_cases), y= total_cases))+
  geom_col(fill = palette)+
  theme_linedraw()+
  theme(axis.text.x = element_text(angle=60, hjust = 1), plot.title = element_text(size = rel(2), hjust = .5)) +
  labs(title = "Top 5 Diseases in California from 2001 to 2019",
       x = "Disease",
       y= "Cases")
```
*We can see the top 5 diseases are Campylobacteriosis, Salmonellosis, Coccidioidomycosis, Giardiasis, Shigellosis.*



3. What county has the most total disease cases?
```{r}
disease_data %>%
  filter(county!="California", sex=="Total") %>%
  group_by(county) %>%
  summarise(sum_cases = sum(cases)) %>%
  arrange(desc(sum_cases))
```
```{r}
disease_data %>%
  filter(county == "Los Angeles", sex == "Total") %>%
  summarise(total_cases = sum(cases))
```
*Los Angeles county has the highest number of cases: 89,165.*

4. County with least total disease cases?
```{r}
disease_data %>%
  filter(county!="California", sex=="Total") %>%
  group_by(county) %>%
  summarise(sum_cases = sum(cases)) %>%
  arrange(sum_cases)
```
```{r}
disease_data %>%
  filter(county == "Alpine", sex == "Total") %>%
  summarise(total_cases = sum(cases))
```
*Alpine county only has 2 total cases!*

Looking at the smallest and greatest populations
```{r}
pop_highest <- disease_data %>%
  filter(county!="California", sex=="Total") %>%
  group_by(county) %>%
  summarise(mean_population_lowest = mean(population)) %>%
  arrange(mean_population_lowest) %>% 
  head(n=1)
pop_highest
```
```{r}
pop_lowest <- disease_data %>%
  filter(county!="California", sex=="Total") %>%
  group_by(county) %>%
  summarise(mean_population_highest = mean(population)) %>%
  arrange(mean_population_highest) %>% 
  tail(n=1)
pop_lowest
```
*The county that has the highest case numbers also has the highest population throughout the years that data was collected. The county with the lowest case numbers also has the lowest average population throughout the years that data was collected.* 

5. Does population affect case numbers for all of the counties? Not just LA and Alpine?

```{r}
cases_overall <- disease_data %>%
  select(county, population, sex, cases) %>%
  filter(county!="California", sex=="Total") %>%
  group_by(county) %>%
  summarise(total_cases = sum(cases)) %>%
  arrange(desc(total_cases))
cases_overall
```
```{r}
disease_data %>%
  ggplot(aes(x= population, y= cases))+
  geom_point(size = .1) +
  theme(plot.title = element_text(hjust = .5)) + theme_get() +
  labs(title = "Population vs Cases from 2001 to 2019",
       x = "Population",
       y= "Cases")

```
*It seems like when there is a higher population there is a higher number of cases and vice versa, which is expected. More on this later!!*

6. Are males or females more impacted by disease in California
```{r}
disease_data %>%
  filter(county!="California") %>%
  group_by(sex) %>%
  summarise(sum_cases = sum(cases, na.rm=T)) %>%
  arrange(desc(sum_cases))
```

```{r}
palette2 <- paletteer_d("wesanderson::Moonrise2")
```

```{r}
disease_data %>% 
  filter(county=="California", sex != "Total") %>%
  group_by(sex) %>% 
  summarise(sum_cases = sum(cases, na.rm=T)) %>%
  ggplot(aes(x=sex, y=sum_cases, fill = sex)) +
  geom_col()+
  scale_fill_manual(values = palette2) +
    labs(title="Male vs Female Disease Prevalence",
       x="Sex",
       y="Number of Cases")
```
*Males are more effected than females.* 

7. How has infectious disease prevalence changed over time in California?
```{r}
disease_data %>% 
  filter(county=="California", sex=="Total") %>%
  group_by(year) %>% 
  summarise(sum_cases = sum(cases)) %>% 
  ggplot(aes(x=year, y=sum_cases))+
  geom_line()+
  geom_point(shape=1, color="Blue")+
  labs(title="Disease Prevalence in California from 2001-2019",
       x="Year",
       y="Number of Cases")
```

*Increased! But has the population has also increased?*

```{r}
disease_data %>% 
  filter(county=="California", sex=="Total") %>%
  group_by(year) %>% 
  summarise(sum_cases = sum(population)) %>% 
  ggplot(aes(x=year, y=sum_cases))+
  geom_line()+
  geom_point(shape=1, color="Blue")+
  labs(title="Population in California from 2001-2019",
       x="Year",
       y="Population Number")
```

*Yes, so we cannot say increase in disease is due to disease factors alone, there are also way more people! We can take the proportion of disease to the population over time to see if there was an actual disease incidence increase.*

```{r}
scaling_pop_cases <- disease_data %>%
  filter(county!="California", sex=="Total") %>%
  group_by(year) %>%
  summarise(sum_population = sum(population),
            sum_cases = sum(cases)) %>%
  mutate(scaling_cases_and_population = sum_cases/sum_population) 
scaling_pop_cases
```

```{r}
scaling_pop_cases %>% 
  ggplot(aes(x=year,y=scaling_cases_and_population))+
  geom_line() + geom_point(color="Blue")+
  labs(title="Percent of Population of California Infected From 2001-2019",
       x="Year",
       y="Percent of Population Infected")
```
*We see that the percent of the population infected over time increases, so we can say that the increase in disease prevalence isn't just because there are more people to infect.*

## Shiny App

```{r}
dat <- disease_data %>% filter(sex=="Total") %>% select(!sex)

ui <- dashboardPage(skin="purple",
  dashboardHeader(title = "Disease Abundance"),
  dashboardSidebar(disable = T),
  dashboardBody(
  fluidRow(
  box(title = "Plot Options", width = 3,
  selectInput("fill", "Select Disease", choices=unique(disease_data$disease)),
                hr(),
      helpText("The data represent cases with an estimated illness onset date from 2001 through the last year indicated from California Confidential Morbidity Reports and/or Laboratory Reports")
  ),
  
  box(title = "What is the Disease Incidence per California County?", width = 9,
  plotOutput("plot", width = "1200px", height = "500px", click="plot_click"),
  verbatimTextOutput("info")
  ),
  
  box(title="What Disease is the Most Prevalent in California from 2001-2019?", width=7,
      plotOutput("plot2", width="1800px", height="600px")
    )
  ) 
  )
  ) 

server <- function(input, output, session) { 
  
  output$plot <- renderPlot({
  disease_data %>% 
  filter(county!="California", disease==input$fill) %>% 
  ggplot(aes_string(x = "county", y="cases", fill="county")) + 
  geom_col()+
  theme_light(base_size = 18)+
  labs (x="County", y="Cases", fill="Disease")+
  theme(axis.text.x = element_text(angle = 60, hjust = 1), legend.position="none")
  })
  
  output$plot2 <-renderPlot({
    disease_data %>%
  filter(sex == "Total", county != "California") %>%
  ggplot(aes(x=reorder(disease, cases), y= log10(cases)))+
  geom_col(fill="purple") +
      theme_light(base_size=18)+
      labs (x="Disease", y="Log10 Cases", fill="Cases")+
  theme(axis.text.y = element_text(size=12), legend.position="none") + coord_flip()
  })

  
   output$info <- renderText({
    paste0("x=", input$plot_click$x, "\ny=", input$plot_click$y)
})
  session$onSessionEnded(stopApp)
  
} 
shinyApp(ui, server)
```
